---
title: "Code Violations GIS Visualization"
author: "Rory Tikalsky & Skyler Halbritter"
date: "March 9, 2016"
output: html_document
---


```
setwd()
```

##Visualizing Using Point Mapping
####Map Basics
```{r}
library(ggmap)
syracuse <- get_map(location="Syracuse NY", zoom = 13, color="bw")
syr.map <- ggmap(syracuse, extent = "device")
```


***

#####This code was originally used to create the dataframes from the original file. Because of the length of geocoding, this was not run for the R Markdown. Instead, outputs were saved in a dataframe and converted to new CSV files.

****
dat2 <- read.csv("new data.csv")

**Open Violations**
open.dat <- dat2[dat2$Complaint.Status == "Open", ]

**Open Bed Bug Violations**
open.bedbugs <- open.dat[open.dat$Complaint.Type == "Bed Bugs",]
open.bedbugs["fulladdress"] <- NA
open.bedbugs$fulladdress <- paste(open.bedbugs$Address, "Syracuse, NY", sep=" ")
long.lat.bugs <- geocode(open.bedbugs$fulladdress)
open.bedbugs <- cbind(open.bedbugs, long.lat.bugs)

**All Bed Bug Violations**
bed.bugs.all <- dat2[dat2$Complaint.Type == "Bed Bugs",]
bed.bugs.all["fulladdress"] <- NA
bed.bugs.all$fulladdress <- paste(bed.bugs.all$Address, "Syracuse, NY", sep=" ")
long.lat.all.bugs <- geocode(bed.bugs.all$fulladdress)
bed.bugs.all <- cbind(bed.bugs.all, long.lat.all.bugs)


**Unsafe conditions**
unsafe.dat <- dat2[dat2$Complaint.Type == "Unsafe Conditions",]
unsafe.dat["fulladdress"] <- NA
unsafe.dat$fulladdress <- paste(unsafe.dat$Address, "Syracuse, NY", sep=" ")
long.lat.unsafe <- geocode(unsafe.dat$fulladdress)
unsafe.dat <- cbind(unsafe.dat, long.lat.unsafe)

**Add color vector to unsafe conditions**
col.vec <- rep( NA, nrow(unsafe.dat) )
col.vec <- ifelse( unsafe.dat$Violation.Status == "Open", "orange", "blue")

***





####Load new dataset (and Subset and Geocode -- see above code)
```{r}
open.bedbugs <- read.csv("openbedbugs.csv")
bed.bugs.all <- read.csv("bedbugsall.csv")
unsafe.dat <- read.csv("unsafe.csv")

#add color vector to unsafe conditions
col.vec <- rep( NA, nrow(unsafe.dat) )
col.vec <- ifelse( unsafe.dat$Violation.Status == "Open", "orange", "blue")

```

####Load Multiplot Function
This function is not in the ggmap package but is widely used for it.
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


####Plot All 3 Maps
```{r}

multiplot(
syr.map + geom_point(data=open.bedbugs, aes(x=lon, y=lat), size=3, col="orange", alpha=.8) +
  ggtitle("Open Bed Bug Violations") +
  theme(plot.title=element_text(family="Times", face="bold", size=12)),
syr.map + geom_point(data=bed.bugs.all, aes(x=lon, y=lat), size=3, col="orange", alpha=.4) +
  ggtitle("All Bed Bug Violations 2012-Present") +
  theme(plot.title=element_text(family="Times", face="bold", size=12)),
syr.map + geom_point(data=unsafe.dat, aes(x=lon, y=lat), size=3, col=col.vec, alpha=.4) +
  ggtitle("All Unsafe Condition Violations 2012-Present") +
  theme(plot.title=element_text(family="Times", face="bold", size=12)),
layout=matrix(c(1,2,3,3), nrow=2, byrow=TRUE)
)
```


------

##Visualizing Using Choropleth Mapping
```{r}
require(maptools)
require(sp)

# get your data sets
syr <- readShapePoly( fn="01-05-2015", proj4string=CRS("+proj=longlat +datum=WGS84") )

violations <- read.csv("Violation Report with lot ID.csv")

#cast as numeric
syr$SBL <- as.character(syr$SBL)
violations$Identifier <- as.character(violations$Identifier)

#merge them
syr@data <- data.frame(syr@data, violations[match(syr@data[, "SBL"], violations[, "Identifier"]), ])

###################################################################
############### Trash Violations in University Area ###############
###################################################################

#subset for university area
univ <- syr[which(syr@data$Nhood=="Westcott"| syr@data$Nhood=="University Neighborhood" | syr@data$Nhood=="University Hill" | syr@data$Nhood=="Near Eastside" | syr@data$Nhood=="South Campus" | syr@data$Nhood=="Outer Comstock"), ]

#create color vector of open violations, closed violations and no violations
col.vec <- ifelse( univ@data$Complaint.Type=="Trash/Debris-Private, Occ" & univ@data$Violation.Status=="Open", "steel blue", "gray80")
col.vec <- ifelse( univ@data$Complaint.Type=="Trash/Debris-Private, Occ" & univ@data$Violation.Status=="Closed", "dark orange", col.vec)
col.vec[ is.na(col.vec) ] <- "gray80"

#plot and create legend
plot( univ, col=col.vec, border=FALSE, main="Trash Violations near Syracuse University (2012-2015)", cex.main=1.4)
legend(x=626459, y=1110819, legend= c("Open Violations", "Closed Violations", "No Violations"), fill=c("steel blue","dark orange", "gray80"), bty="n", cex=.8)


#########################################################################
############## Overgrowth Violations in University Area #################
#########################################################################

#create color vector of open violations, closed violations and no violations
col.vec <- ifelse( univ@data$Complaint.Type=="Overgrowth: Private, Occ" & univ@data$Violation.Status=="Open", "steel blue", "gray80")
col.vec <- ifelse( univ@data$Complaint.Type=="Overgrowth: Private, Occ" & univ@data$Violation.Status=="Closed", "dark orange", col.vec)
col.vec[ is.na(col.vec) ] <- "gray80"

#plot and create legend
plot( univ, col=col.vec, border=FALSE, main="Overgrowth Violations Near Syracuse University (2012-2015)", cex.main=1.4)
legend(x=626459, y=1110819, legend= c("Open Violations", "Closed Violations", "No Violations"), fill=c("steel blue","dark orange", "gray80"), bty="n", cex=.8)
```
