---
title: "Lab07_Advanced Graphics"
author: "Skyler Halbritter"
date: "October 26, 2015"
output: html_document
---
```{r}
library(Lahman)
data("Salaries")
```

####1.	Calculate each team's budget from 2002 using the tapply() function.
```{r}
sal.2002 <- Salaries [Salaries$yearID=="2002", ]
sal.2002.names <- merge( sal.2002, Teams, by.x=c("teamID", "yearID", "lgID"), by.y=c("teamID", "yearID", "lgID"))
sal.2002 <- tapply( X=sal.2002.names$salary, INDEX=sal.2002.names$name, sum, na.rm=T )
sal.2002.dat <- data.frame( budget=as.numeric(sal.2002), name=names(sal.2002))
sal.2002 <- merge(sal.2002.dat, Teams, by.x="name", by.y="name" )
sals.2002 <- sal.2002[sal.2002$yearID =="2002",]
sals.2002 <- sals.2002[ order( sals.2002$budget, decreasing=T ), ]
sals.2002[c("name", "budget")]
```

***

####2.	Subset the results by American and National League. Create separate graphs for each league. Include the league average budget in the appropriate spot on the graph.
####3.	Use each team's total budget instead of the % of league budget.
####4. Label everything appropriately. 

**National League Subset**
```{r}
nl.2002 <- sals.2002[sals.2002$lgID == "NL",]
nl.ave <- mean(nl.2002$budget)
nl.ave.dat <- data.frame(budget=nl.ave, name="National League Average")
nl.sal.2002 <- merge(nl.2002, nl.ave.dat, all.x = T, all.y=T)
nl.2002 <- nl.sal.2002[ order( nl.sal.2002$budget, decreasing=T ), ]
```
**National League Plot**
```{r}
plot( nl.2002$budget, 17:1, type="n",
      xlim=c(-50000000,150000000), ylim=c(-2,20), 
      bty="n", ylab="", yaxt="n", xlab="", xaxt="n"
)
rect (xleft=0, ybottom=9.5, xright=125000000, ytop=10.5, col="gray94", border="gray94")
segments( x0=0, y0=0, y1=17, col="gray", lty=3 )
segments( x0=25000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=50000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=75000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=100000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=125000000, y0=0, y1=17, col="gray", lty=3 )
title(main="2002 Budgets of National League Teams"  )
segments( x0=0, x1=nl.2002$budget, y0=17:1, col="lightgray", lty=1 )
text( nl.2002$budget, 17:1, (paste("$", (format(nl.2002$budget , big.mark=",")), sep="")), pos=4, cex=0.6, col="lightskyblue3" )
text( c(25000000, 50000000, 75000000, 100000000, 125000000), -1, c("$25M","$50M","$75M","$100M","$125M"), col="gray43", cex=0.6 )
text( -1, 17:1, nl.2002$name, col = "gray43", pos=2, cex=0.6)
points(nl.2002$budget, 17:1, pch=19, cex=.8, col="lightskyblue3")
```

***
**American League Subset**
```{r}
al.2002 <- sals.2002[sals.2002$lgID == "AL",]
al.ave <- mean(al.2002$budget)
al.ave.dat <- data.frame(budget=al.ave, name="American League Average")
al.sal.2002 <- merge(al.2002, al.ave.dat, all.x = T, all.y=T)
al.2002 <- al.sal.2002[ order( al.sal.2002$budget, decreasing=T ), ]
```
**American League Plot**
```{r}
plot( al.2002$budget, 15:1, type="n",
      xlim=c(-50000000,150000000), ylim=c(-2,18), 
      bty="n", ylab="", yaxt="n", xlab="", xaxt="n"
)
rect (xleft=0, ybottom=8.5, xright=125000000, ytop=9.5, col="gray94", border="gray94")
segments( x0=0, y0=0, y1=15, col="gray", lty=3 )
segments( x0=25000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=50000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=75000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=100000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=125000000, y0=0, y1=15, col="gray", lty=3 )
title(main="2002 Budgets of American League Teams"  )
segments( x0=0, x1=al.2002$budget, y0=15:1, col="lightgray", lty=1 )
text( al.2002$budget, 15:1, (paste("$", (format(al.2002$budget , big.mark=",")), sep="")), pos=4, cex=0.6, col="indianred" )
text( c(25000000, 50000000, 75000000, 100000000, 125000000), -1, c("$25M","$50M","$75M","$100M","$125M"), col="gray43", cex=0.6 )
text( -1, 15:1, al.2002$name, col = "gray43", pos=2, cex=0.6)
points(al.2002$budget, 15:1, pch=19, cex=.8, col="indianred")
```

***
**National and American League Comparison**
```{r, echo=FALSE}
par( mfrow=c(2,1), mar=c(0,0,0,0 ))
plot( nl.2002$budget, 17:1, type="n",
      xlim=c(-50000000,150000000), ylim=c(-2,20), 
      bty="n", ylab="", yaxt="n", xlab="", xaxt="n"
)
rect (xleft=0, ybottom=9.5, xright=125000000, ytop=10.5, col="gray94", border="gray94")
segments( x0=0, y0=0, y1=17, col="gray", lty=3 )
segments( x0=25000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=50000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=75000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=100000000, y0=0, y1=17, col="gray", lty=3 )
segments( x0=125000000, y0=0, y1=17, col="gray", lty=3 )
title(main="2002 Budgets of National League Teams", line=-1  )
segments( x0=0, x1=nl.2002$budget, y0=17:1, col="lightgray", lty=1 )
text( nl.2002$budget, 17:1, (paste("$", (format(nl.2002$budget , big.mark=",")), sep="")), pos=4, cex=0.6, col="lightskyblue3" )
text( c(25000000, 50000000, 75000000, 100000000, 125000000), -1, c("$25M","$50M","$75M","$100M","$125M"), col="gray43", cex=0.6 )
text( -1, 17:1, nl.2002$name, col = "gray43", pos=2, cex=0.6)
points(nl.2002$budget, 17:1, pch=19, cex=.8, col="lightskyblue3")
plot( al.2002$budget, 15:1, type="n",
      xlim=c(-50000000,150000000), ylim=c(-5,15), 
      bty="n", ylab="", yaxt="n", xlab="", xaxt="n"
)
title(main="2002 Budgets of American League Teams", line=-10)
rect (xleft=0, ybottom=8.5, xright=125000000, ytop=9.5, col="gray94", border="gray94")
segments( x0=0, y0=0, y1=15, col="gray", lty=3 )
segments( x0=25000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=50000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=75000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=100000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=125000000, y0=0, y1=15, col="gray", lty=3 )
segments( x0=0, x1=al.2002$budget, y0=15:1, col="lightgray", lty=1 )
text( al.2002$budget, 15:1, (paste("$", (format(al.2002$budget , big.mark=",")), sep="")), pos=4, cex=0.6, col="indianred" )
text( -1, 15:1, al.2002$name, col = "gray43", pos=2, cex=0.6)
points(al.2002$budget, 15:1, pch=19, cex=.8, col="indianred")
```
